## The Next Step: kableExtra

kableExtra is a library that makes it much easier to build on 
`knitr::kable` tables without having to muck around with CSS. 

```{r}
#| label: kableExtra-more-options
mtcars |> 
  head() |> 
  kable(
    col.names = c("Miles/(US) Gallon", "Number of Cylinders",
                  "Displacement (cu. in.)", "Gross Horsepower", "Rear Axle Ratio",
                  "Weight (1000 lbs.)", "1/4 Mile Time", 
                  "Engine (0=V-shaped, 1=Straight",
                  "Transmission: (0=automatic, 1=manual)",
                  "Number of Forward Gears",
                  "Number of Carburetors"),
    caption = "Motor Trend Car Road Tests",
    align = "r",
    digits=c(1, 0, 0, 0, 2, 1, 1, 0, 0, 0, 0)
  )
```

`knitr::kable` and `kableExtra::kable` are basically identical functions,
as seen above. `kableExtra` also includes the `kbl` function, which
produces kable tables but has better auto-completion and help info 
for the various options `knitr` has under the hood, especially for
TeX tables.

### Getting Started

The fast way to clean up a `kable` is to add one of the functions
from `kableExtra`:
  
  - `kable_classic`
- `kable_classic_2`
- `kable_material`
- `kable_material_dark`
- `kable_minimal`
- `kable_paper`


Examples (with code and some columns omited for brevity):
  
  ```{r}
#| label: kable-func

# creates a basic kable to pipe into the syling options:
kableFunc <- function(caption){
  mtcars |> 
    select(mpg:wt) |> 
    head() |> 
    kbl(
      caption = caption,
      align = "r",
    )
}
```



```{r}
#| label: kable-classic
kableFunc(caption = "kable_classic") |> 
  kable_classic()
```

```{r}
#| label: kable-classic-2

kableFunc(caption = "kable_classic_2") |> 
  kable_classic_2()
```

```{r}
#| label: kable-material

kableFunc(caption = "kable_material") |> 
  kable_material()
```

```{r}
#| label: kable-material-dark

kableFunc(caption = "kable_material_dark") |> 
  kable_material_dark()
```

```{r}
#| label: kable-minimal

kableFunc(caption = "kable_minimal") |> 
  kable_minimal()
```

```{r}
#| label: kable-paper

kableFunc(caption = "kable_paper") |> 
  kable_paper()
```

All of these options provide pretty good tables. 
When further customization is desired, enter: `kable_styling`.

### Getting Stylish