## Knitr::kable, Ol' Reliable

### The Basics

The `kable` function from the Knitr library is always a good place 
to start when putting together a table.
(Note that kable must be specified with `knitr::kable` since we've
  loaded kableExtra)
  
```{r}
#| label: kable-basic
mtcars |> 
  head() |> 
  knitr::kable()
```

Well that's not too exciting. It's not much different than just 
  printing the dataframe. It just has bold column headers. 
  
  
### Adding some detail

Luckily `kable` has some options that we can invoke.
  For example, detailed column names.

```{r}
#| label: kable-col_names
mtcars |> 
  head() |> 
  knitr::kable(
    col.names = c("Miles/(US) Gallon", "Number of Cylinders",
                  "Displacement (cu. in.)", "Gross Horsepower", "Rear Axle Ratio",
                  "Weight (1000 lbs.)", "1/4 Mile Time", 
                  "Engine (0=V-shaped, 1=Straight",
                  "Transmission: (0=automatic, 1=manual)",
                  "Number of Forward Gears",
                  "Number of Carburetors")
  )
```

That's a little better. At least we know what the columns are, 
  though it's a little cramped.
  
### More Options

  Some other things we can do are:
  
  - Add row names: `row.names=c("Vector", "of", "names")`
    - mtcars already has rownames, so none need to be added here.
  - Add a caption: `caption = "Caption"`
  - Align text: `align = "c"`
    - This can be set table-wide or by column. The string will be repeated
  if there are more columns.
  - Specify the digits, also table-wide or by column.
  
```{r}
#| label: knitr-more-options
mtcars |> 
  head() |> 
  knitr::kable(
    col.names = c("Miles/(US) Gallon", "Number of Cylinders",
                  "Displacement (cu. in.)", "Gross Horsepower", "Rear Axle Ratio",
                  "Weight (1000 lbs.)", "1/4 Mile Time", 
                  "Engine (0=V-shaped, 1=Straight",
                  "Transmission: (0=automatic, 1=manual)",
                  "Number of Forward Gears",
                  "Number of Carburetors"),
    caption = "Motor Trend Car Road Tests",
    align = "r",
    digits=c(1, 0, 0, 0, 2, 1, 1, 0, 0, 0, 0)
  )
```

`knitr::kable` is capable of further formatting options, but they tend
  to require writing custom CSS and passing that through options
  that are not very friendly to work with. See 
  [Yihui's site](https://bookdown.org/yihui/rmarkdown-cookbook/kable.html#customize-html-tables)
                 for more details on those options. 
                 Fortunately there are other tools available that let us stay within
                 the realm of R programming to further modify `knitr::kable` tables.