---
title: "Nice Looking Tables in Quarto"
subtitle: "In HTML"
author: "Tyler Clark"
format: html
toc: true
---

# Making Tables in Quarto

There are several ways to make tables in Quarto, with 
  different pros and cons to them.
  
  
## Why Tables?

One could just print the raw data into the document, like so:

```{r}
#| label: mtcars-basic
head(mtcars)
```

But that's not very friendly or professional, is it? 
  How does someone evaluate what the data is saying?
  Why would they even read your report if you're just spitting out data? 
  That is where the tools available in R and Quarto come in. 


## Tools of the Trade

The libraries that are used in this document are:

- The tidyverse: `library(tidyverse)`
- Knitr: `library(knitr)`
- KableExtra: `library(kableExtra)`
- gt: `library(gt)`
- DataTables: `library(DT)`
- flextable: `library(flextable)`
- huxtable: `library(huxtable)`
- More (`rmarkdown::paged_table`, `reactable`, `formattable`, `pixiedust`) may come later

```{r}
#| label: load-libraries
#| message: false
#| warning: false

library(tidyverse)
library(knitr)
library(kableExtra)
library(gt)
library(flextable)
library(huxtable)
library(DT)
library(reactable)
library(formattable)
library(pixiedust)

```


```{r}
#| label: knitr-section
#| child: "Knitr.qmd"
```
  
```{r}
#| label: kableExtra-section
#| child: "kableExtra.qmd"
```


